Mitigating Low Audio Quality by Domain Transfer for Better ASR

### 2020-04-15 2020-tsd-brno-denoise/denoise.tex
redaction from DKl

Automatic speech recognition is generally approached by means of machine
learning: a statistical model is inferred based on some training data. Its
application bases on the assumption that whatever was learned about the training
data will also apply for the input data. It is clear then, that if the input
data have different properties than the training data, the basic assumption is
violated and the inference performance will suffer.
Also if the training data themselves are damaged in a way that makes the desired
information harder to extract, the learning is going to be less successful.

@@

I am conducting an experiment to make the input data similar to the training
data instead of the more usual other way around because 1) it is not easy to
make the volunteer annotators transcribe acoustically flawed data, 2) a
success would not only benefit speech recognition results but also listening
experience and 3) the model could stay leaner.

What acoustic defects are actually present? Since the corpus was digitized from
magnetophone tapes of spontaneous talks by a single speaker recorded in various
amateur conditions and with consumer devices, we can broadly distinguist these
types of interefering signal:
\begin{enumerate}
\item{additive noise -- hum or hiss,}
\item{
    stationary interference like screeching added by low-quality magnetophone
    parts or the erasing head signal,
}
\item{non-stationary interference like background speech or door slams,}
\item{
    room echo or ill-equalized microphone boosting or cutting certain
    frequencies,
}
\item{non-linear distortion of the magnetophone}
\item{speed fluctuations.}
\end{enumerate}

The individual interference types affect each other and can occur several times
in several stages. For instance, echo is like a convolution with a certain
sample vector. It is followed by non-linear distortion and after that
convolution again with a different vector in the imperfect magnetophone
circuits. Additive noise comes on top of that. Another batch of distortions
comes during the playback. Clearly, modeling, let alone reverting these
interferences is a difficult task.

@@

For curiosity, I looked at the distances of the files among each other and
compared it to that between other familiar sounds. For instance, two adjacent
segments of the same speaker from a meeting of the Czech parliament has a
distance of 1.4. Different speakers have a distance of 6.5. One of these
segments compared to a black metal song chorus has the distance of 519. On the
other hand, the median distance among the Spoken corpus of Karel Makoň, where
there is one single speaker, amounts to 55.9.

The experiment presented below is conducted on three clusters: 1) A cluster
of the cleanest recordings, used as the destination domain, 2) a cluster of
overdriven recordings and 3) one of those recorded with low tape speed. The
latter recordings are the hardest to understand also for humans. All clusters
have a maximum internal distance of 25.

>>>>

Automatic speech recognition is generally approached by means of machine
learning: a statistical model is inferred based on some training data. Its
application is based on the assumption that whatever was learned about the training
data will also apply to the input data. It is clear then, that if the input
data have different properties than the training data, the basic assumption is
violated and the inference performance will suffer.
Also, if the training data themselves are damaged in a way that makes the desired
information harder to extract, the learning is going to be less successful.

@@

I am conducting an experiment to make the input data similar to the training
data instead of the more usual other way around because 1) it is not easy to
make the volunteer annotators transcribe acoustically flawed data, 2) not only
speech recognition would benefit from successful denoising, the listening
experience would improve too, and 3) the model could stay leaner.

What acoustic defects are actually present? Since the corpus was digitized from
magnetophone tapes of spontaneous talks of a single speaker recorded in various
amateur conditions and with consumer devices, we can broadly distinguist these
types of quality degradation:
\begin{enumerate}
\item{additive noise -- hum or hiss,}
\item{
    stationary interference like screeching added by low-quality magnetophone
    parts or the erasing head signal,
}
\item{non-stationary interference like background speech or door slams,}
\item{
    room echo or ill-equalized microphone boosting or cutting certain
    frequencies,
}
\item{non-linear distortion of the magnetophone}
\item{speed fluctuations.}
\end{enumerate}

The individual interference types affect each other and can occur several times
in several stages. For instance, the echo is essentially convolution with a certain vector of
amplitude values, known as the impulse response. It even might not be
constant if there are moving objects in the room or if the speaker or
microphone is moving. This convolution is followed by a non-linear
distortion in the microphone and its preamplifier and convolved again with
another impulse response, this time representing a frequency and phase
response of recorder's amplifiers. Finally it is non-linearly distorted in
the process of making a magnetic recording on the tape. Another batch of distortions
comes during the playback. Clearly, modeling, let alone reverting these
interferences is a difficult task.

@@

For curiosity, I looked at the distances of the files among each other and
compared it to that between other familiar sounds. For instance, two adjacent
segments of the same speaker from a meeting of the Czech parliament has a
distance of 1.4. Different speakers have a distance of 6.5. One of these
segments compared to a black metal song chorus has the distance of 519. On the
other hand, the median distance among the Spoken corpus of Karel Makoň, where
there is a single speaker, amounts to 55.9.

The experiment presented below is conducted on three clusters:
\begin{enumerate}
\item{A cluster of the cleanest recordings, used as the destination domain,}
\item{a cluster of overdriven recordings and}
\item{a cluster of recordings taken with low tape speed.}
\end{enumerate}
Recordings in cluster 3 are the hardest to understand even for humans. All these
clusters have a maximum internal distance of 25.
